---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
library(brfinance)
```

# brfinance 📊🇧🇷

[![CRAN Status](https://www.r-pkg.org/badges/version/brfinance)](https://cran.r-project.org/package=brfinance)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![R-CMD-check](https://github.com/efram2/brfinance/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/efram2/brfinance/actions/workflows/R-CMD-check.yaml)
[![Downloads](https://cranlogs.r-pkg.org/badges/grand-total/brfinance)](https://cran.r-project.org/package=brfinance)
[![Downloads](https://cranlogs.r-pkg.org/badges/brfinance)](https://cran.r-project.org/package=brfinance)
[![GitHub stars](https://img.shields.io/github/stars/efram2/brfinance.svg)](https://github.com/efram2/brfinance/stargazers)
[![GitHub forks](https://img.shields.io/github/forks/efram2/brfinance.svg)](https://github.com/efram2/brfinance/network)

**brfinance** puts Brazil's official macroeconomic data — interest rates, inflation, exchange rates, unemployment, GDP — and a set of financial-math calculators behind a single, consistent R interface. One function call gets you a clean `data.frame`; one more gets you a ready-to-share `ggplot2` chart.

If your work touches **Brazilian interest rates**, **inflation**, **exchange rates**, or plain **financial math** (present/future value, NPV, IRR, amortization...), this package saves you from hand-rolling BCB API calls every time.

> **Looking for stock/equity data instead?** brfinance focuses on macro series (rates, inflation, GDP) from the Central Bank and IBGE. For individual B3 stocks, portfolios, and Markowitz-style analysis, see the companion package **[{brstocks}](https://github.com/efram2/brstocks)**.

---

## Why brfinance?

With **brfinance**, you can:

- Pull **official Brazilian macro indicators** (BCB & IBGE) with one function call
- Get the **Ibovespa index** and compare it against rates/inflation on the same chart
- Run standard **financial calculations** (NPV, IRR, PV, FV, PMT, amortization schedules, and more)
- Plot any indicator with **one line of code** — consistent styling across the whole package
- Get output in **Portuguese or English**, your choice, without changing how you call the function

No manual API requests. No scraping. No reshaping raw JSON by hand — brfinance does that once, so you don't have to do it every time.

---

# Installation

```r
install.packages("brfinance")

# Or the development version, straight from GitHub
install.packages("devtools")
devtools::install_github("efram2/brfinance")

library(brfinance)
```

# 🚀 Quick Start

Every `get_*()` function returns a plain `data.frame` with `date` and `value` columns (plus a couple of extras where it makes sense, like `get_ipca_from_target()` below) — so once you've learned one, you've learned the pattern for all of them.

**Inflation (IPCA)**

```{r fig.ipca, fig.width=7, fig.height=4}
plot_inflation_rate(
  get_inflation_rate("2020", "2024")
)
```

**SELIC interest rate**

```{r fig.selic, fig.width=7, fig.height=4}
plot_selic_rate(
  get_selic_rate("2020", "2024")
)
```

**Unemployment rate**

```{r fig.unemp, fig.width=7, fig.height=4}
plot_unemployment(
  get_unemployment("2019", "2024")
)
```

**Ibovespa index**

Want equity-market context alongside your macro series? `get_ibovespa()` pulls the daily Ibovespa close (via Yahoo Finance) in the same `date`/`value` shape as everything else, so it drops straight into `plot_ibovespa()` or into a comparison chart with any other indicator below.

```{r fig.ibov, fig.width=7, fig.height=4}
plot_ibovespa(
  get_ibovespa("2020-01-01", "2024-12-31")
)
```

*(Remember: this is the benchmark index only. For individual stocks, portfolio construction, or an efficient-frontier simulator, that's what [{brstocks}](https://github.com/efram2/brstocks) is for.)*

**IPCA vs. the inflation target**

A common question in Brazilian macro analysis: is inflation running above or below the Central Bank's target? `get_ipca_from_target()` answers it directly — it returns the 12-month accumulated IPCA, the CMN's official target for that year, and the gap between them, all in one `data.frame`.

```{r fig.ipca-target}
ipca_gap <- get_ipca_from_target("2015", "2024")
head(ipca_gap)
```

**Compare multiple indicators in one chart**

You can compare different economic indicators three ways:

1. Raw values, on their original scale
2. Indexed series (base = first observation = 100)
3. Percentage change from the first observation

```{r fig.compare, fig.width=8, fig.height=5}
plot_series_comparison(
  data_list = list(
    "SELIC" = get_selic_rate("2020", "2024"),
    "IPCA"  = get_inflation_rate("2020", "2024"),
    "Unemployment" = get_unemployment("2020", "2024")
  ),
  y_vars = rep("value", 3),
  date_vars = rep("date", 3),
  scale_type = "index",
  title = "Brazilian Economic Indicators",
  subtitle = "Indexed comparison (base = first observation)"
)
```

**...or on two separate axes, keeping each series' native scale**

When the series you're comparing have very different units — a rate in % vs. an exchange rate in R\$ — indexing them can hide what actually matters. Set `dual_axis = TRUE` (only for exactly two series) to plot each on its own y-axis instead:

```{r fig.dual, fig.width=8, fig.height=5}
plot_series_comparison(
  data_list = list(
    "SELIC" = get_selic_rate("2020", "2024"),
    "Exchange Rate" = get_exchange_rate("2020-01-01", "2024-12-31")
  ),
  y_vars = c("value", "value"),
  date_vars = c("date", "date"),
  dual_axis = TRUE,
  title = "SELIC vs. USD/BRL Exchange Rate"
)
```

**Financial calculators**

```{r calculators}
# Net Present Value
calc_npv(rate = 0.1, cashflows = c(-1000, 300, 400, 500))

# Internal Rate of Return
calc_irr(cashflows = c(-1000, 300, 400, 500))

# Loan payment
calc_pmt(pv = 10000, rate = 0.02, n = 24)

# Full amortization schedule for a loan
head(calc_amortization_schedule(pv = 10000, rate = 0.01, n = 12))
```

# Function Reference

## Data retrieval functions (`get_*`)

* `get_inflation_rate()` — Brazil's official inflation rate (IPCA), monthly
* `get_ipca_from_target()` — 12-month accumulated IPCA vs. the CMN's official inflation target, plus the gap between them
* `get_selic_rate()` — Daily SELIC interest rate, annualized (base 252)
* `get_cdi_rate()` — Daily CDI interbank rate, annualized
* `get_exchange_rate()` — USD/BRL exchange rate
* `get_gdp_growth()` — GDP growth rates
* `get_unemployment()` — Unemployment rate (PNAD Contínua)
* `get_ibovespa()` — Ibovespa index closing level (via Yahoo Finance)
* `browse_series()` / `get_series_info()` — Explore the full catalog of Central Bank (SGS) series available, beyond the ones wrapped above

All `get_*` functions accept flexible date inputs: `"YYYY"`, `"YYYY-MM"`, or `"YYYY-MM-DD"`. Requests spanning more than 10 years are chunked and stitched together automatically, since that's the Central Bank API's own limit per request — not something you need to work around yourself.

---

## Plotting functions (`plot_*`)

* `plot_inflation_rate()` — Plots IPCA inflation over time
* `plot_selic_rate()` — Plots the SELIC interest rate
* `plot_cdi_rate()` — Plots the CDI interest rate
* `plot_exchange_rate()` — Plots exchange rate time series
* `plot_unemployment()` — Plots unemployment rate time series
* `plot_ibovespa()` — Plots the Ibovespa index
* `plot_series_comparison()` — Compares multiple economic indicators in one chart, indexed/raw/percent-change, or on a dual axis (`dual_axis = TRUE`)

Every `plot_*()` function shares the same visual language: a financial-market blue for benchmark rates (SELIC, CDI, exchange rate, Ibovespa) and a market red for cost/risk indicators (inflation, unemployment), so charts from different functions read consistently side by side.

---

## Financial calculators (`calc_*`)

* `calc_present_value()` — Present Value
* `calc_future_value()` / `calc_compound_interest()` — Future Value / Compound Interest (same formula, two names for two common ways people search for it)
* `calc_future_value_ext()` — Future Value with periodic payments
* `calc_simple_interest()` — Simple interest
* `calc_effective_rate()` / `calc_nominal_rate()` — Convert between nominal and effective rates
* `calc_pv_annuity()` / `calc_fv_annuity()` — Present/future value of an annuity
* `calc_pmt()` — Loan payment amount
* `calc_rate()` — Solve for the interest rate per period
* `calc_nper()` — Solve for the number of periods
* `calc_npv()` — Net Present Value
* `calc_irr()` — Internal Rate of Return
* `calc_amortization_schedule()` — Full loan amortization schedule
* `calc_continuous_compounding()` / `calc_pv_continuous()` — Continuous compounding, future and present value
* `rule_of_72()` / `rule_of_114()` — Quick estimate of years to double / triple an investment

# Language Support

Every `get_*()` function accepts a `language` argument:

* `language = "eng"` (default) — English variable labels
* `language = "pt"` — Portuguese variable labels

Column names themselves (`date`, `value`, ...) stay the same either way, so switching `language` never breaks a pipeline built on top of a `get_*()` call — it only changes the descriptive labels attached to the data (visible via `labelled::var_label()`).

# Data sources

All data comes from **official Brazilian institutions**:

* Central Bank of Brazil (BCB / SGS)
* IBGE (SIDRA / PNAD Contínua)
* Yahoo Finance, for the Ibovespa index (`get_ibovespa()`)

# Related package

For individual B3 stocks, portfolio simulation, and an interactive Markowitz efficient-frontier dashboard, check out **[{brstocks}](https://github.com/efram2/brstocks)** — brfinance's sister package, focused on the equity side of Brazilian markets.

# Contribution

Suggestions, feature requests, and pull requests are welcome!
