---
title: "README"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
library(brfinance)
```

# brfinance 📊🇧🇷

[![CRAN Status](https://www.r-pkg.org/badges/version/brfinance)](https://cran.r-project.org/package=brfinance)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![R-CMD-check](https://github.com/efram2/brfinance/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/efram2/brfinance/actions/workflows/R-CMD-check.yaml)
[![Downloads](https://cranlogs.r-pkg.org/badges/grand-total/brfinance)](https://cran.r-project.org/package=brfinance)
[![Downloads](https://cranlogs.r-pkg.org/badges/brfinance)](https://cran.r-project.org/package=brfinance)
[![GitHub stars](https://img.shields.io/github/stars/efram2/brfinance.svg)](https://github.com/efram2/brfinance/stargazers)
[![GitHub forks](https://img.shields.io/github/forks/efram2/brfinance.svg)](https://github.com/efram2/brfinance/network)

**brfinance** is an R package that brings together **official Brazilian macroeconomic data** and **practical financial calculators** in a simple, ready-to-use interface.

If you work with **Brazilian data**, **interest rates**, **inflation**, or **financial math**, this package is for you.

---

## Why brfinance?

With **brfinance**, you can:

- Access **official Brazilian macroeconomic indicators** (BCB & IBGE)
- Run **financial calculations** (NPV, IRR, PV, FV, PMT, rates)
- Plot key indicators with **one line of code**
- Use **Portuguese or English** outputs seamlessly

No APIs. No scraping. No manual cleaning.

---

# Installation

```r
install.packages("brfinance")

# Or development version from GitHub
install.packages("devtools")
devtools::install_github("efram2/brfinance")

library(brfinance)

```
# 🚀 Quick Start

Inflation (IPCA)

```{r fig.ipca, fig.width=7, fig.height=4}
plot_inflation_rate(
  get_inflation_rate("2020", "2024")
)
```

SELIC interest rate

```{r fig.selic, fig.width=7, fig.height=4}
plot_selic_rate(
  get_selic_rate(2020, 2024)
)
```

Unemployment rate

```{r fig.unemp, fig.width=7, fig.height=4}
plot_unemployment(
  get_unemployment("2019", "2024")
)
```

Compare multiple indicators in one chart

* Want to analyze how different economic indicators evolve together?
* Use *plot_series_comparison()* to compare multiple time series in a single, clean visualization.

```{r fig.compare, fig.width=8, fig.height=5}
plot_series_comparison(
  data_list = list(
    "SELIC" = get_selic_rate(2020, 2024),
    "IPCA"  = get_inflation_rate("2020", "2024"),
    "Unemployment" = get_unemployment("2020", "2024")
  ),
  y_vars = rep("value", 3),
  date_vars = rep("date", 3),
  scale_type = "index",
  title = "Brazilian Economic Indicators",
  subtitle = "Indexed comparison (base = first observation)"
)
```

## Financial Calculators (minimalista)

```r
# Net Present Value
calc_npv(0.1, c(-1000, 300, 400, 500))

# Internal Rate of Return
calc_irr(c(-1000, 300, 400, 500))

# Loan payment
calc_pmt(rate = 0.02, n = 24, pv = 10000)
```
# Function Reference

## Data retrieval functions (`get_*`)

* `get_inflation_rate()` → Retrieves Brazil’s official inflation rate (IPCA)
* `get_selic_rate()` → Retrieves the SELIC interest rate
* `get_cdi_rate()` → Retrieves the CDI interest rate
* `get_exchange_rate()` → Retrieves official exchange rates
* `get_gdp_growth()` → Retrieves GDP growth rates
* `get_unemployment()` → Retrieves unemployment rate from PNAD Contínua
* `get_sgs_series()` → Retrieves any Central Bank (SGS) time series by code

All `get_*` functions support flexible date inputs:
`"YYYY"`, `"YYYY-MM"`, or `"YYYY-MM-DD"`.

---

## Plotting functions (`plot_*`)

* `plot_inflation_rate()` → Plots IPCA inflation over time
* `plot_selic_rate()` → Plots the SELIC interest rate
* `plot_cdi_rate()` → Plots the CDI interest rate
* `plot_exchange_rate()` → Plots exchange rate time series
* `plot_gdp_growth()` → Plots GDP growth over time
* `plot_unemployment()` → Plots unemployment rate time series
* `plot_series_comparison()` → Compares multiple economic indicators in one chart

---

## Financial calculators (`calc_*`)

* `calc_pv()` → Present Value
* `calc_fv()` → Future Value
* `calc_npv()` → Net Present Value
* `calc_irr()` → Internal Rate of Return
* `calc_pmt()` → Loan payment calculation
* `calc_rate()` → Interest rate calculation
* `calc_nper()` → Number of periods
* `calc_compound_interest()` → Compound interest calculation
* `calc_continuous_interest()` → Continuous compounding
* `calc_rule72()` → Rule of 72 approximation
* `calc_rule114()` → Rule of 114 approximation

# Language Support

All main functions support bilingual output:

* language = "eng" (default): Returns English column names and labels
* language = "pt": Returns Portuguese column names and labels

# Data sources

All data come from *official Brazilian institutions:*

* Central Bank of Brazil (BCB / SGS)
* IBGE (SIDRA / PNAD Contínua)

# Contribution

Suggestions, feature requests, and pull requests are welcome!
